Approximating the Optimal State Abstraction

### 2015-10-06 state_abstraction.tex
added notes from dave-ellis meeting from 10/6

% --- SECTION: Outline of H2R Talk ---
\newpage
\section{Outline of H2R Talk}

@@

\vspace{4mm}
\elem{For example:}
In this lab, we like OO-MDPs a lot. There are other possible representations one might consider to collapse our representation of the real world. Objects seem like a reasonable choice given how object-oriented the human experience seems to be.

@@

% - Subsection: Unified Theory Paper -
\midline
\subsection{Unified Theory Paper}
Michael, Tom Walsh, Lihong Li wrote a paper called ``Towards a Unified Theory of State Abstraction for MDPs''~\cite{li2006towards} that surveys these approaches, and puts them into a single framework.

@@

\elem{Example time!}

First: Sample Complexity (e.g. PAC-MDP means $\epsilon-$optimalilty with polynomial Sample Complexity in the relevant parameters, w/ probability $1-\delta$).
Second: Grid example that goes from NxM to N.

Q: But what's the elephant in the room with all of this?
A: Finding states with {\it exactly} the same Q values, or optimal action and action value, is rare!
A: Furthermore, we need to know something about the optimal behavior in advance!

\midline
\subsection{New Proposal}

An area I'm interested in investigating: lets relax this assumption of {\bf strict equality} and look at some notion of similarity/approximate equality, and then investigate what is preserved in the resulting policy.

This will let us back off and see what might be learned online in a lifelong reinforcement learning setting. That is, can investigate learning one of these $\phi$ functions online.

Tackles:
\begin{itemize}
\item Lifelong RL
\item Online learning
\item State abstraction
\item Macroaction/Option learning
\end{itemize}

% --- SECTION: Introduction ---
\newpage
\section{Introduction}

@@

>>>>

% --- SECTION: Motivation ---
\newpage
\section{Background}

@@

\vspace{4mm}
\elem{For example:}
Historically, we'ved relied on OO-MDPs a lot. There are other possible representations one might consider to collapse our representation of the real world. Objects seem like a reasonable choice given how object-oriented the human experience seems to be.

@@

% Subsection: Unified Theory Paper
\midline
\subsection{Unified Theory Paper}
Michael, Tom Walsh, Lihong Li wrote a paper called ``Towards a Unified Theory of State Abstraction for MDPs''~\cite{li2006towards} that surveys these approaches, and puts them into a single framework.

@@

% --- SECTION: Motivation ---
\newpage
\section{Motivation}

% Subsection: Arbitrary reduction in Sample Complexity
\subsection{Result 1: Arbitrary reduction in Sample Complexity}
First: Grid example that goes from NxM to N, (always go up, no down action).

\subsection{Result 2: With a perfect abstractor, consider the exploration problem}

\subsection{Result 3: Collapsing on {\it just state variables} is insufficient to preserve any kind of optimality}

\subsection{Elephant in the Room}
Q: But what's the elephant in the room with all of this?
A: Finding states with {\it exactly} the same Q values, or optimal action and action value, is rare!
A: Furthermore, we need to know something about the optimal behavior in advance!

\midline

% --- SECTION: New Proposal ---
\newpage
\section{New Proposal: Approximate Abstraction}

For each of the following three cases:
% Model Equivalence
\begin{equation}
 \phi(s_1) = \phi(s_2) \rightarrow \forall_{s,a} : T(s \mid s_i, a) = T(s \mid s_j, a)
\end{equation}

% Optimal Q Function Equivalence
\begin{equation}
 \phi(s_1) = \phi(s_2) \rightarrow \forall_a : Q^*(s_i, a) = Q^*(s_j, a)
\end{equation}

% Optimal action and value Equivalence
\begin{equation}
 \phi(s_1) = \phi(s_2) \rightarrow \argmax_a Q^*(s_i, a) = \argmax_a Q^*(s_j, a) \wedge \max_a Q^*(s_i, a) = \max_a Q^*(s_j, a)
\end{equation}

Relax the equality condition, instead consider the approximate case, e.g.:
% Model Similarity
\begin{equation}
 \phi(s_1) = \phi(s_2) \rightarrow \forall_{s,a} : |T(s \mid s_i, a) - T(s \mid s_j, a)| \leq \epsilon
\end{equation}

\elem{The Question:} Can anything be said about the optimality of the policy $\pi_{\phi_{T,\epsilon}}$ in the ground MDP? Are we within some factor of $\epsilon$ away from being optimal?

% --- SECTION: Temporal Abstraction ---
\newpage
\section{New Proposal: Temporal Abstraction}

In light of Result 3, that state abstraction needs to be defined with respect to some sort of properties about actions, suppose a set of temporally extended actions are included in the action set.

Now, abstractions become temporally extended.

Should be some nice results there.

@@

Future Work:
\begin{itemize}
\item Learning $\phi$
\item Connection to AMDPs (e.g. consider $\Omega : \langle S, A, R, T, \gamma \rangle \mapsto \langle S', A', R', T', \gamma' \rangle$
\item Connection to teaching.
\end{itemize}
